Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2023-04-26 paper-correlations.tex
Style and typos fixes

\section{The scan approach.}
\label{sec:scan}
All algorithms described above are executed on the whole input and hence require
a lot of memory. The scan approach deals with memory limitations at the cost of
higher execution time. In this approach we select a list of predefined
directions and cut all one dimensional slices from the input along those
directions. Then for each direction we use an appropriate algorithm in
\cref{sec:map} to calculate a correlation function just for that slice. When all
slices are processed we calculate the result using the formula:
\begin{equation*}
  F^{\bm{d}}(r) = \frac{\sum\limits_i F^{\bm{d}}_i(r) N_i(r)}{\sum\limits_i N_i(r)}
\end{equation*}
Here $F^{\bm{d}}(r)$ is a correlation function computed with respect to a
direction $\bm{d}$, $F^{\bm{d}}_i(r)$ is a ``partial'' correlation
function calculated for a $i$-th slice and $N_i(r)$ is a total number of trials
for a correlation length $r$ and slice $i$.

@@

Another function which does not use this approach but is in the
\code{Directional} module for historical reasons is the pore size function
$P(r)$. Remember that pore size function is a probability distribution
function. In practice it is better represented as a histogram of pore size
fractions, so all bars of the histogram sum to 1. In our approach we calculate
this histogram using distance transform. For each point $\bm{x}$ distance
transform $\mathcal{D}(\bm{x})$ is defined as
\begin{equation*}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \min\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right.
\end{equation*}
Distance transform has many implementations \textcolor{red}{cite something}. Our
algorithm for calculating $P(r)$ is then:
\begin{algorithmic}[1]
  \Procedure{$P$}{$A, nbins$}
    \State $D \gets \mathcal{D}(A)$ \Comment{Apply $\mathcal{D}$ to each element of $A$}
    \State $D' = \{\forall x \in D: x \ne 0\}$
    \State Bin values in $D'$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

\section{Verifying our algorithms}
\label{sec:verification}
Our algorithms are verified by comparing their results with known closed-form
expressions for correlation functions for some sort of input data. One kind of
well-studied data is overlapping n-balls with fixed radius and centers generated
by Poisson point process. Here we consider two-dimensional and three-dimensional
cases.

@@

\appendix
\section{Algorithm of computation of lineal-phase function}
\label{linpathalg}
The algorithm below calculates $L_2^{(phase)}$ correlation function for
one-dimensional array $a$. The result is an array of $L_2$ values for
correlation lengths from $0$ to $maxlen-1$. If $periodic$ is true (resp. false)
then periodic (resp. zero padding) boundary conditions are used. The entry point of the
algorithm is the function L2. The efficiency of the proposed algorithm is $O(n)$
where $n$ is the length of the array.

>>>>

\section{The scan approach.}
\label{sec:scan}
All algorithms described above are executed on the whole input and hence require
a lot of memory. The scan approach deals with memory limitations at the cost of
higher execution time. In this approach we select a list of predefined
directions and cut all one-dimensional slices from the input along those
directions. Then for each direction we use an appropriate algorithm in
\cref{sec:map} to calculate a correlation function just for that slice. When all
slices are processed we calculate the result using the formula:
\begin{equation*}
  F^{\bm{d}}(r) = \frac{\sum\limits_i F^{\bm{d}}_i(r) N_i(r)}{\sum\limits_i N_i(r)}
\end{equation*}
Here $F^{\bm{d}}(r)$ is a correlation function computed with respect to a
direction $\bm{d}$, $F^{\bm{d}}_i(r)$ is a ``partial'' correlation
function calculated for a $i$-th slice and $N_i(r)$ is a total number of trials
for a correlation length $r$ and slice $i$.

@@

Another function which does not use this approach but is in the
\code{Directional} module for historical reasons is the pore size function
$P(r)$. Remember that pore size function is a probability distribution
function. In practice it is better represented as a histogram of pore sizes
found in the input. In our approach we calculate this histogram using distance
transform. For each point $\bm{x}$ distance transform $\mathcal{D}(\bm{x})$ is
defined as
\begin{equation*}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \min\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right.
\end{equation*}
Distance transform has many implementations \textcolor{red}{cite something}. Our
algorithm for calculating $P(r)$ is then:
\begin{algorithmic}[1]
  \Procedure{$P$}{$A, nbins$}
    \State $D \gets \mathcal{D}(A)$ \Comment{Apply $\mathcal{D}$ to each element of $A$}
    \State $D' = \{ x \in D: x \ne 0\}$
    \Comment Remove zeros
    \State Bin values in $D'$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

\section{Verification of our algorithms}
\label{sec:verification}
Our algorithms are verified by comparing their results with known closed-form
expressions for correlation functions for some sort of input data. One kind of
well-studied data is overlapping n-balls with fixed radius and centers generated
by Poisson point process. Here we consider two-dimensional and three-dimensional
cases.

@@

\appendix
\section{Algorithm of computation of lineal-path function}
\label{linpathalg}
The algorithm below calculates $L_2^{(phase)}$ correlation function for
one-dimensional array $a$. The result is an array of $L_2$ values for
correlation lengths from $0$ to $maxlen-1$. If $periodic$ is true (resp. false)
then periodic (resp. zero padding) boundary conditions are used. The entry point of the
algorithm is the function L2. The efficiency of the proposed algorithm is $O(n)$
where $n$ is the length of the array.
